---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r setoptions, echo = TRUE, cache = TRUE}
```

## Loading required package

library(ggplot2)
library(hms)

## Loading and preprocessing the data

```{r loadingData}
acData <- read.csv(file = "activity.csv", header = TRUE)
```

## What is mean total number of steps taken per day?

### calculate total of steps taken per day

```{r total}
tSteps <- aggregate(steps ~ date, acData, FUN = sum)
```

### Histogram of total steps taken per day

```{r histogram}
hist(tSteps$steps,
      main = "Total Steps Taken per Day",
      xlab = "Number of Steps")
```

### mean and median of the total steps taken per day

```{r mean and median per day}
meanSteps <- mean(tSteps$steps, na.rm = TRUE)
medianSteps <- median(tSteps$steps, na.rm = TRUE)
```

## What is the average daily activity pattern?

### average daily activity pattern

```{r average daily activity pattern}
meansin <- aggregate(steps ~ interval, acData, mean)
```

### time series of daily average activity

```{r time series}
ggplot(data = meansin, aes(x = interval, y = steps)) +
     geom_line() +
     ggtitle("Average Daily Activity Pattern") +
     xlab("5-minute Interval") +
     ylab("Average Steps Numbers") +
     theme(plot.title = element_text(hjust = 0.5))
```

### max steps by interval

```{r max steps per interval}
maxin <- meansin[which.max(meansin$steps), ]
```

## Imputing missing values

### total number of missing values

```{r total NA}
mvalue <- is.na(acData$steps)
table(mvalue)
```

### imputing missing values

```{r imputing NA}
filled_acData <- transform(acData,
     steps = ifelse(is.na(acData$steps),
     meansin$steps[match(acData$interval,
     meansin$interval)],
     acData$steps))
```

### daily steps after imputing missing data

```{r total after imputing}
filled_interval <- aggregate(steps ~ date, filled_acData, FUN = sum)
```

### Histogram of daily steps after imputing missing data

```{r Hist after imputing}
hist(filled_interval$steps,
      main = "Daily Steps (Imputed Number)",
      xlab = "Number of Steps")
```

### mean and median of the total steps taken per day after imputing

```{r mean and median after imputing}
imMeanSteps <- mean(filled_interval$steps, na.rm = TRUE)
imMedianSteps <- median(filled_interval$steps, na.rm = TRUE)
```

## Are there differences in activity patterns between weekdays and weekends?

### differences in new factors

```{r differences in new factors}
difMean = imMeanSteps - meanSteps
difMedian = imMedianSteps - medianSteps
difTotal = sum(filled_interval$steps) - sum(tSteps$steps)
```

### new factor variable in the dataset with two levels

```{r factor with two levels}
Day_type <- function(date) {
         day <- weekdays(date)
         if (day %in% c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday'))
         return ("weekeday")
         else if (day %in% c('Saturday', 'Sunday'))
         return ("weekend")
         else
         stop ("Invalid Date Format.")
 }
filled_acData$date <- as.Date(filled_acData$date)
filled_acData$day <- sapply(filled_acData$date, FUN = Day_type)
 ```
 
### average steps taken by weekdays and weekend
```{r average per day_type}
Dailymean <- aggregate(steps ~ interval + day, filled_acData, mean)
```

### comparison plot of activity pattern between weekdays and weekend
```{r gplot}
ggplot(data = Dailymean, aes(x = interval, y = steps)) +
         geom_line() +
         facet_grid(day ~ .) +
         ggtitle("Average Daily Activity Pattern") +
         xlab("5-minute Interval") +
         ylab("Average Steps Numbers") +
         theme(plot.title = element_text(hjust = 0.5))
```
